---
output: github_document
bibliography: inst/ref.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- knit with rmarkdown::render("README.Rmd", output_format = "md_document") -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = FALSE
)
```

# Individual Variance Detection

<!-- badges: start -->
[![R-CMD-check](https://github.com/ph-rast/ivd/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ph-rast/ivd/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/ph-rast/consistentlyBetter-CanvasApp/graph/badge.svg?token=V6PK2J7AL9)](https://codecov.io/gh/ph-rast/consistentlyBetter-CanvasApp)
<!-- badges: end -->

SSranef is an R package for random effects selection that corresponds to the models described in @rodriguez2021. Specifically, `ivd_alpha()` fits a random intercepts model with a spike-and-slab prior on the random effects and `ivd_beta()` fits a model with both random intercepts and random slopes, with a spike-and-slab prior on the random effects for the slope. The function `ivd_mv()` fits a multivariate mixed-effects models for two outcomes and places a spike-and-slab prior on the random slope for each outcome.

This package is repurposed from the orginal verision at [josue-rodriguez/SSranef](https://www.github.com/josue-rodriguez/SSranef)

## Installation


This package can be installed with 

``` r
# install.packages("devtools")
devtools::install_github("consistentlybetter/ivd")
```

## Example


```{r example}
library(ivd)

d <- mlmRev::Hsb82
```

## Alpha model
```{r}
d$y <- c(scale(d$mAch))

alpha <- ivd_alpha(y = d$y, unit = d$school)

posterior_summary(alpha, ci = 0.90, digits = 2)
```

```{r}
ranef_summary(alpha, ci = 0.95, digits = 2)
```

```{r}
caterpillar_plot(alpha, legend = FALSE)
```

```{r}
pip_plot(alpha, legend = FALSE)
```

## Beta model
```{r}
beta <- ivd_beta(y = d$y, X = d$ses, unit = d$school)
posterior_summary(beta, digits = 2)
```

```{r}
ranef_summary(beta, digits = 2)
```

```{r}
caterpillar_plot(beta, legend = FALSE)
```

```{r}
pip_plot(beta, legend = FALSE)
```

```{r include=FALSE, eval=FALSE}
library(lme4)
summary(lmer(y ~ 1 + (1|school), data = d))
```

## Multivariate model

```{r}
mv_data <- gen_mv_data(5, 5)
str(mv_data)

mv_model <- ivd_mv(Y = cbind(mv_data$y1, mv_data$y2),
                        X = mv_data$x,
                        unit = mv_data$id,
                        burnin = 100,
                        iter = 500,
                        chains = 4)

posterior_summary(mv_model)
```


## Priors

Priors can be passed on to either of the `ivd` functions through a named list and using JAGS code, e.g., 

```{r}
# change prior for mean intercept
priors <- list(alpha = "alpha ~ dt(0, 1, 3)",
               # for each jth unit, change prior probability of inclusion
               gamma = "gamma[j] ~ dbern(0.75)") 

fit <- ivd_alpha(y = d$y, unit = d$school, priors = priors)
ranef_summary(fit)
```

## Building on top of SSranef models

The code for each model can also be extracted to make more extensive modifications or build more complex models

```{r}
jags_model_text <- fit$model_text
cat(jags_model_text)
```

# References
